---
title: "Homework 2"
author: Danyang Gui
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```


## Problem 1

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )

```


Read precipitation data!

```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)
  
precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
   
```

Now combine annual precipitation.

```{r}
month_df =
  tibble(
    month  = 1:12,
    month_name = month.name
  )

precip_df= 
     bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
     
```



This dataset contains information from the Mr. TrashWheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final datasets. Additional data sheets include month precipitation data. 




## Problem 2



```{r}
subway_df = 
  read_csv ("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(-division, -exit_only, -staffing, -staff_hours, -ada_notes, -free_crossover, 
         -north_south_street, -east_west_street, -corner, -entrance_latitude,       -entrance_longitude,
         -station_location, -entrance_location) %>%
  mutate(entry = recode (entry, 'YES' = TRUE , 'NO' = FALSE ))
```


This data set is about subway system in NYC. I imported the csv file into R and briefly cleaned the variable names with janitor::clean_names. Then, I excluded variables such as staffing, entrance latitude, and entrance longitude,etc. There are `r nrow(subway_df) ` rows and `r ncol(subway_df) ` columns in the existing dataframe now. After the first round of cleaning, variables retained are `r names(subway_df)`. Thus far, the data is not yet tidy.  


```{r}
subway_df_clean =
  distinct(subway_df, station_name, line, ada) %>% 
  filter (ada == TRUE) %>% 
  nrow()
```

There are 465 distinct stations. 84 of them are ADA compliant. 



```{r}

proportion_denominator = 
  select (subway_df, entry, vending) %>% 
  filter (vending == "NO") %>% 
  nrow()

proportion_numerator = 
  select (subway_df, entry, vending) %>% 
  filter (entry == TRUE, vending == "NO") %>% 
  nrow()


proportion_numerator / proportion_denominator

```


37.70492% of the entrances/exits without vending allow entrance. 


```{r}

subway_df_route = 
  mutate_at(subway_df, vars(route8:route11), as.character) %>%
  pivot_longer(
    route1:route11,
    names_to = "route_name",
    values_to = "route_number"
  )

```

from route 8 to route 11 are in the format of double, we need to first convert them to character. 

```{r}
filter(subway_df_route, route_number == "A") %>% 
  select(route_number, line, station_name) %>% 
  distinct(route_number, line, station_name)


filter(subway_df_route, route_number == "A", ada == "TRUE") %>% 
  select(route_number, ada, line, station_name) %>%  
  distinct(route_number, ada, line, station_name)

```

60 distinct stations serve the A train. 

Of the 60 distinct stations that serve the A train, 17 are ADA compliant. 








## Problem 3


Firstly,let's deal with the data in pols-month.csv. 
Import and clean the pols-months data, break up the variable mon into integer variables year, month, and day; replace month number with month name

```{r}

pols_month_df = 
          read_csv("./data/pols-month.csv") %>% 
          janitor::clean_names() %>% 
          separate(mon, c("year", "month", "day"), sep = "([-])") %>% 
          mutate(month = as.numeric(month),
                 day = as.numeric(day),
                 year = as.numeric(year)
                 ) %>% 
          relocate(prez_gop, prez_dem) %>% 
          pivot_longer(
                 prez_gop:prez_dem,
                 names_to = "president",
                 names_prefix = "prez_",
                 values_to = "president_number"
          ) %>% 
          filter(president_number == 1) %>% 
          select(-day) 
          
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

pols_month_df = 
         left_join(pols_month_df, month_df, by = "month") %>% 
         relocate(year, month_name) %>% 
         select(-month) %>% 
         rename(month = month_name)

```




Secondly, we move on to snp.csv data.

```{r}
snp_df = 
  read_csv ("./data/snp.csv") %>%
  janitor::clean_names() %>% 
  separate(date, c("month", "day", "year"), sep = "([/])") %>% 
  mutate(month = as.numeric(month),
         day = as.numeric(day),
         year = as.numeric(year)) %>% 
  select(-day) 
          
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

snp_df = 
         left_join(snp_df, month_df, by = "month") %>% 
         relocate(year, month_name) %>% 
         select(-month) %>% 
         rename(month = month_name)
```



Third, we import and tidy the unemployment data.

```{r}
umemploy_df = 
  read_csv("./data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment_percentage"
    ) %>% 
 mutate(month = recode(month, "jan" = "January", "feb" = "February", "mar" = "March", 
                        "apr" = "April", "may" = "May", "jun" = "June", "jul" = "July",
                        "aug" = "August", "sep" = "September", "oct" = "October",
                        "nov" = "November", "dec" = "December"))
          
```


Now, let's merge them together. 

```{r}
merge_df =
  left_join(pols_month_df, snp_df, by = c("year", "month")) %>%
  left_join(., umemploy_df, by = c("year", "month"))
```


The pols_month_df contains `r nrow(pols_month_df)` rows and `r ncol(pols_month_df)`columns.The range of year is`r range(pull(pols_month_df, year)) `. Variables include `r names(pols_month_df) `.

The snp_df contains `r nrow(snp_df)` rows and `r ncol(snp_df)`columns.The range of year is`r range(pull(snp_df, year)) `. Variables include `r names(snp_df) `.

The umemploy_df contains `r nrow(umemploy_df)` rows and `r ncol(umemploy_df)`columns.The range of year is`r range(pull(umemploy_df, year)) `. Variables include `r names(umemploy_df) `.

The merged dataset contains `r nrow(merge_df)` rows and `r ncol(merge_df)`columns.
The range of year is `r range(pull(merge_df, year)) `. The number of governors, senators, representatives in both republican and democratic are represented afterwards. From which party in that given time the president was elected followed then. The closing values of the S&P stock index and unemployment percentage on the associated month are also shown.



